Plant functional type effects on biomass partitioning

### 2014-12-09 manuscript.Rnw
Some rewriting of opening para

@@

\textbf{The partitioning of biomass in terrestrial vegetation between leaves, stems and roots is a critical component of dynamic global vegetation models (DGVMs). Biomass partioning is consistently shown to be one of the most uncertain and important drivers for global predictions of carbon storage \cite{ise_comparison_2010}, the lifetime of carbon in ecosystems \cite{friend_carbon_2014}, and the CO\textsubscript{2} response of net primary production \cite{dekauwe2014}. Little is known about direct and indirect effects of climate on biomass partitioning, largely due to the lack of a global individual-based biomass dataset. Here, we analyze the largest database on plant biomass to date [cite]. We show that partitioning of biomass to leaves vs. woody stems is strongly dependent on the plant functional type. Evergreen gymnosperms maintain three times more leaf biomass at a given total aboveground biomass than deciduous angiosperms, with evergreen angiosperms taking an intermediate value. We found that leaf mass per area, an indicator of leaf construction cost, explains these differences so that the PFTs converged in the ratio  of leaf area to aboveground biomass. We found weak effects of climate while accounting for PFT effects, suggesting that climate-related differences are captured largely by a shift in PFT. These results can be used to constrain biomass partitioning routines in global dynamic vegetation models.}

% \section{flow}
%
% \begin{itemize}
%
% \item First paragraph
% \item Allocation in models, main uncertainties
% \item Climate effects on partioning : direct (MAT, MAP), indirect (PFT).
% \item Known: gymnosperm vs. angiosperm. Unknown: leaf habit (deciduous vs. evergreen), direct climate, patterns for individual plants.
% \item BAAD : Largest database, other large datasets are stands or only small plants
% \item Found very large LMF differences, but convergent LAR
% \item decompose into MF/AS, MT/AS, find similar differences in MF/AS but none in MT/AS, suggesting convergent AF/AS between PFTs.
% \item Only small direct climate effects, so shift in PFT captures global patterns in biomass partioning.
% \item This has implications for allocation in DGVMs, they can be divided into roughly three categories in terms of biomass allocation (allocation fractions, allometric constraints, LAI optimization). Allocation vs. partitioning: the role of leaf and stem turnover.
%
% \end{itemize}

@@

Previous syntheses on plant biomass relationships has predominantly focused on the universal nature of size-biomass relationships in plants. Little is known about differences between plant functional types (evergreen vs. deciduous, angiosperm vs. gymnosperm). Since most global models parameterize allocation submodels by PFT, it is imperative that we better understand these potential differences.

@@

%------------------------------------------------------------------------------------%
\section*{Methods}

>>>>

% \section{flow}
%
% \begin{itemize}
%
% \item First paragraph
% \item Allocation in models, main uncertainties
% \item Climate effects on partioning : direct (MAT, MAP), indirect (PFT).
% \item Known: gymnosperm vs. angiosperm. Unknown: leaf habit (deciduous vs. evergreen), direct climate, patterns for individual plants.
% \item BAAD : Largest database, other large datasets are stands or only small plants
% \item Found very large LMF differences, but convergent LAR
% \item decompose into MF/AS, MT/AS, find similar differences in MF/AS but none in MT/AS, suggesting convergent AF/AS between PFTs.
% \item Only small direct climate effects, so shift in PFT captures global patterns in biomass partioning.
% \item This has implications for allocation in DGVMs, they can be divided into roughly three categories in terms of biomass allocation (allocation fractions, allometric constraints, LAI optimization). Allocation vs. partitioning: the role of leaf and stem turnover.
% \item Another implication is that, as a first approximation, using LMA is more appropriate than splitting by PFT. Using LMA, together with the assumption of invariant leaf area display, gives biomass partitioning between leaves and stems. This even worked, to some extent, across species.
%
% \end{itemize}

@@

\textbf{The partitioning of biomass in terrestrial vegetation between long-lived stems and short-lived leaves affects the lifetime of carbon in ecosystems \cite{friend_carbon_2014}, model predictions of global carbon storage \cite{ise_comparison_2010}, and the CO\textsubscript{2} response of net primary production \cite{dekauwe2014}. Little is known about direct and indirect effects of climate on biomass partitioning, largely due to the lack of a global individual-based biomass dataset. Here, we analyze a new database on plant biomass fractions, the largest to date [cite]. We show that partitioning of biomass to leaves vs. woody stems is strongly dependent on the plant functional type, and proportional to the leaf mass per area, an important leaf functional trait. As a result, plant functional types converged in the ratio of leaf area to aboveground biomass. We found only modest effects of climate after accounting for PFT, suggesting that climate-related differences in biomass partitioning are captured largely by a shift in PFT. These results form an important constraint on biomass partitioning routines in global dynamic vegetation models.}

@@

% 1. Importance of allocation in global vegetation models, short-lived leaves vs. long-lived stems affect ecosystem-scale turnover of C, as well as productivity (as partitioning to leaves affects LAI). DGVMs split plant diversity into a few plant functional types (PFTs). This begs two questions : how does biomass partitioning dffer between PFTs, and can we relate it to commonly measured leaf traits, to simplify application of DGVMs?

Previous syntheses on plant biomass relationships have predominantly focused on the universal nature of size-biomass relationships in plants. Relatively little is known about differences in biomass partitioning between PFTs(evergreen vs. deciduous, angiosperm vs. gymnosperm).

@@

\section*{Methods}
